---
title: " RMarkdown workshop homework day 1"
author: "Younjung Choi"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float:
      collapsed: false
    number_sections: true
---

# Set-up libraries and data
```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	error = TRUE,
	warning = FALSE,
	dev = c("png", "pdf", "svg"),
	tidy = TRUE
)
```

## Load libraries
```{r load libraries, message=FALSE}
library(tidyverse)
library(pheatmap)
```

## Load data
```{r load data}
load("data/Rmarkdown_data.Rdata")
```

# Top 20 significant genes

## Get names of top 20 genes
```{r top 20 sig genes}
top20_sigOE_genes <- res_tableOE_tb %>% 
  arrange(padj) %>% 	#Arrange rows by padj values
  pull(gene) %>% 		#Extract character vector of ordered genes
  head(n=20)	

```

## normalized counts for top 20 significant genes
```{r normalized counts of top 20 sig genes}
top20_sigOE_norm <- normalized_counts %>%
  filter(gene %in% top20_sigOE_genes)
```

## Gathering the columns to have normalized counts to a single column
```{r}
gathered_top20_sigOE <- top20_sigOE_norm %>%
  gather(colnames(top20_sigOE_norm)[2:9], key = "samplename", value = "normalized_counts")
gathered_top20_sigOE <- inner_join(mov10_meta, gathered_top20_sigOE)

```

## plot using ggplot2
```{r scatter plot}
ggplot(gathered_top20_sigOE) +
  geom_point(aes(x = gene, y = normalized_counts, color = sampletype)) +
  scale_y_log10() +
  xlab("Genes") +
  ylab("log10 Normalized Counts") +
  ggtitle("Top 20 Significant DE Genes") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))
```


# Create a heatmap of the differentially expressed genes

## Extract normalized expression for significant genes from the OE and control samples (2:4 and 7:9)
```{r filter significant genes}
res_tableOE_tb_sig <- res_tableOE_tb %>%
  filter(padj < 0.05)
```


## Return the normalized counts for the significant DE genes
```{r normalized counts of sig genes}
norm_OEsig <- normalized_counts %>% 
  filter(gene %in% res_tableOE_tb_sig$gene) 
```


```{r metadata formatting}
meta <- mov10_meta %>%
column_to_rownames("samplename") %>%
data.frame()
```


## Run pheatmap using the metadata data frame for the annotation
```{r pheatmap}
pheatmap(norm_OEsig[2:9], 
         cluster_rows = T, 
         show_rownames = F,
         annotation = meta, 
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         fontsize_row = 10, 
         height = 20)
```

**conclusion**: example. if i want to use a *italic* link is [here](https://raw.githubusercontent.com/hbctraining/reproducibility-tools/master/activities/Rmd_exercise4_answerkey.rmd).


## Project details
```{r session information}
sessionInfo()
```

